Sophize Knowledge Organization

### 2022-05-27 math_library/sophize_math.tex
For CICM'22 submission 4/X

\subsection{Core concepts}
Sophize uses the following concepts to logically organize theoretical knowledge.

@@

\subsection{Validity of arguments}

@@

\subsection{Belief sets}
\label{sec:bset}

@@

\subsection{Aggregating knowledge}

@@

\subsection{PGM response}
The response of a request to any PGM has the following components:

@@

We have extended the Markdown language - a lightweight markup language for creating formatted text using a plain-text editor - for building a rich narration interface. It allows users to quickly encode and present semantic data, to show the truth status of propositions available in prose, and to effortlessly access proof graphs. The detailed specification of Sophize Markdown is available here. Many of its features are demonstrated at: \url{https://youtu.be/5UYOpQwcjCk}. We highlight some of the salient features here.

@@

\subsection{Status Indicators}

For propositions, a truth-value icon indicating whether or not there is a proof is added next to the link. Clicking on the icon brings up the proof graph as shown in Figure \ref{proof_tree}. Similarly, an icon is added next to a proof generating machine which indicates whether the PGM is active or not. These options can be turned on or off depending on the input and the context.

@@

In formal languages, definitions of terms, statements of propositions, and supporting argument text may be parseable by an external parser. We can convert the externally parsed output into Markdown in such a case, where each term is automatically linked to the appropriate resource. This provides a convenient interface, where the user types in the native language, and the final output automatically allows users to explore all concepts that make up the input statement in depth. Currently, this is demonstrated in the use of Sophize Markdown with the Metamath language.

@@

\subsection{\LaTeX\space Support}

With Sophize Markdown, we can add \LaTeX\space by enclosing it between two \$ signs (or \$\$ for display math).

The following example summarizes the specification:
\begin{quote}
\begin{verbatim} 
Einstein's most famous equation is $E=mc^2$ but it is his handwritten theory of 
happiness is what fetched $1.3 million recently. While it may not fetch many \$s, I 
like the field equations more: $$G_{\mu \nu }+\Lambda g_{\mu \nu }=\kappa T_{\mu \nu }$$
\end{verbatim}
\end{quote}

The above Markdown code is rendered as:
\begin{mdframed}
Einstein's most famous equation is $E=mc^2$ but it is his handwritten theory of happiness is
what fetched \$1.3 million recently. While it may not fetch many \$s, I like the field
equations more: $$G_{\mu \nu }+\Lambda g_{\mu \nu }=\kappa T_{\mu \nu }$$
\end{mdframed}

\subsection{Sophize Markdown benefits}

@@

\section{Other Engineering challenges}
\todo{copyright/access management}

\todo{ storing and reading data in disk/GitHub like repo }
\subsubsection{Other Notes}
\todo{What criteria are met based on the 2013 report}

>>>>

\subsection{Core Concepts}
Sophize uses the following concepts to logically organize theoretical knowledge.

@@

\subsection{Validity of Arguments}

@@

\subsection{Belief Sets}
\label{sec:bset}

@@

\subsection{Aggregating Knowledge}

@@

\subsection{PGM Response}
The response of a request to any PGM has the following components:

@@

We have extended the Markdown language - a lightweight markup language for creating formatted text using a plain-text editor - for building a rich narration interface. It allows users to quickly encode and present semantic data, to show the truth status of propositions available in prose, and to effortlessly access proof graphs.

With Sophize Markdown, we can easily add \LaTeX\space by enclosing it between two \$ signs (or \$\$ for display math). The detailed specification of Sophize Markdown is available here. Many of its features are demonstrated at: \url{https://youtu.be/5UYOpQwcjCk}. We highlight the prominent features here.

@@

\subsection{Truth Status}

Sophize markdown adds an icon next to each proposition link. This icon indicates whether the proposition has been proven or dis-proven (or both) in the currently selected belief set using a check-mark or a cross. Clicking on the icon brings up the proof graph as shown in Figure \ref{proof_tree}. If the users switches from one belief set to another, the icon is updated based on the truth values in the new belief set but other parts of the article are unchanged.

Similarly, an icon is added next to a proof generating machine which indicates whether the PGM is active or not.

@@

In formal languages, definitions of terms, statements of propositions, and supporting argument text may be parse-able by an external parser. We can convert the externally parsed output into Markdown in such a case, where each term is automatically linked to the appropriate resource. This provides a convenient interface, where the user types in the native language, and the final output automatically allows users to explore all concepts that make up the input statement in depth. Currently, this is demonstrated in the use of Sophize Markdown with the Metamath language.

@@

\section{Building a Modern Mathematics Library}
The argument and proof graphs in help us build a network of information that spans across a wide variety of sources (including live results from PGMs). Using different belief sets, users can view mathematical knowledge arising from different theoretical foundations and choose what they consider to be reliable. With Sophize Markdown, users can easily embed semantic information, thus making it convenient to explore the network of knowledge that is being built. However, a few more practical issues need to be overcome to build a modern scalable library with reliable information.

\subsection{Semantic Data Extraction}
Utilizing the full range of features offered by the Sophize platform requires the availability of semantic information. However, extracting semantic information from unstructured or semi-structured literature is a difficult problem. This problem can be seen as composed of several sub-problems:
\begin{itemize}
\item Finding definitions and propositions in prose.
\item Associating the appropriate entity with a term used in prose for quick lookup.
\item Identifying arguments and along with premises and conclusion.
\item \label{prob:dup} Identifying duplicate definitions and propositions used in different sources.
\end{itemize}

Solving these problems would need manual effort (e.g. crowd-sourcing) combined with automated extraction tools. Recent progress in machine learning and natural language processing techniques does make this problem seem tractable.

\subsubsection{}

\subsubsection{Connecting Across Datasets}
The sub-problem 4 mentioned above is crucial to help make connections across datasets. For example, once we have identified that the `Jensen's Inequaility' in a published paper is same as the Planetmath proposition with the same name, Planetmath's proof is made available to the reader of the published paper.

\subsection{Search Interface}

\subsection{Data Management}
Sophize organizes

\subsection{Collaboration}

\todo{copyright/access management}

\todo{ storing and reading data in disk/GitHub like repo }

@@

\subsubsection{Other Notes}
\todo{What criteria are met based on the 2013 report}
